\artus{} -- Framework for Event-based Data Analysis in High Energy Physics

### 2015-10-23 Core/doc/latex/paper.tex
Paper: Further changes after discussion

\usepackage{booktabs}
\usepackage{hepnames}
\usepackage{hepunits}
\usepackage{url}
\usepackage{palatino}
\usepackage[hidelinks,bookmarks=true,plainpages=false]{hyperref}
\usepackage[colorinlistoftodos,prependcaption]{todonotes}

@@

\begin{frontmatter}
\title{\artus -- Framework for Event-based Data Analysis in High Energy Physics}

\author{Fabio Colombo}
\author{Raphael Friese}
\author{Dominik Haitz}
\author{Thomas Hauth}
\author{Thomas M\"uller}
\author{G\"unter Quast}
\author{Georg Sieber}
\address{Karlsruhe Institute of Technology, Wolfgang-Gaede-Stra\ss{}e 1, D-76131 Karlsruhe}

@@

\section{Typical Workflow of High Energy Physics Analyses}
\label{section_artus_analysis_workflow}
% This section should contain a short description of a HEP/CMS-typical workflow.
% Especially common problems and challenges should be highlighted.
% From these problems, the design requirements for Artus are derived.
The workflow of the experiments collecting data at the Large Hadron Collider (LHC) at CERN is taken as an example to illustrate the organization and needs of a modern High Energy Physics (HEP) analysis.
Still, equal or very similar requirements are also present for HEP experiments not located at the LHC.

@@

The reconstruction of the raw data coming from the experiments is performed on large computing centers organized in the WLCG\footnote{\url{http://wlcg-public.web.cern.ch}}, the Worldwide LHC Computing Grid.
The resulting datasets contain event information and high-level physics objects and are of $\mathcal O(10-100)$~TB in size.
\todo[inline]{move file size numbers to Figure \ref{figure_analysis_steps}?}

The data which are relevant for a specific analysis are selected in a step usually referred to as \textit{skimming}.
The output of the skimming process is an analysis-dependent file, usually in an n-tuple format, and small enough to be stored on local file servers, often owned and operated by a single institute.
The size of these files ranges from $\mathcal O(10)$~GB to few~TB.

@@

\todo[inline]{I would not put the remaining part of this chapter here.
It should be either part of the requirements or part of the summary what Artus does achieve; but in this technical part in between it feels misplaced
-> Should we put this part into a dedicated section directly after this one?
I think the idea was to explain how certain design decisions, e.g.\ pipelines, achieve the \artus design goals.
Therefore it should not be too far from the technical descriptions.
}
This structure extends the established concepts of existing frameworks like FWLite from CMSSW to the following four main advantages:
\begin{description}
\item[Modular structure:] Analysis steps are implemented in processors, breaking down the complete analysis in well-defined and easily understandable and writable parts.
\item[Re-utilization of code:] Processors can be shared among different analysis groups.
Every user contributes to the testing of existing code and improves the confidence in this code.
\item[Configurability:] Each module uses settings that are read from the the configuration.
Each pipeline is mapped to a separate sub-set of the configuration.
\item[High performance:] The concept of pipelines allows to analyze the data with different configurations at the same time, while the data is read in only once and common processing steps are shared in the global pipeline.
Therefore, the cost-intensive input operations are reduced to a minimum.
\end{description}

@@

%time HiggsToTauTauAnalysis.py @HiggsAnalysis/KITHiggsToTauTau/data/ArtusWrapperConfigs/Sync13TeV.cfg -i HiggsAnalysis/KITHiggsToTauTau/data/Samples/NAF_sample_DYJetsToLLM50_RunIISpring15DR74_Asympt25ns_13TeV_MINIAOD_amcatnloFXFX-pythia8_recent.txt -f 10
\begin{table}[h!]
\centering
\begin{tabular}{lrr}
\toprule
\textbf{Pipelines}                                & \textbf{Runtime / \minute} & \textbf{Event Rate / \reciprocal\second} \\ \midrule
%0                                                & 5:53                             & 1296 \\
 0                                                & 5:12                             & 1470 \\ \midrule
%\em                                              & 6:27                             & 1181 \\
 1 ($\Pe\Pe$)                                     & 6:31                             & 1174 \\
%1 ($\Pe\Pgt$)                                    & 7:03                             & 1084 \\
 1 ($\Pe\Pgt$)                                    & 7:26                             & 1028 \\
%1 ($\Pgm\Pgt$)                                   & 7:14                             & 1054 \\
 1 ($\Pgm\Pgt$)                                   & 7:14                             & 1052 \\
%1 ($\Pgt\Pgt$)                                   & 6:14                             & 1224 \\
 1 ($\Pgt\Pgt$)                                   & 6:07                             & 1247 \\ \midrule
%2 ($\Pgm\Pgt$, $\Pe\Pgt$)                        & 8:33                             & 893 \\
 2 ($\Pgm\Pgt$, $\Pe\Pgt$)                        & 8:32                             & 896 \\
%3 ($\Pgm\Pgt$, $\Pe\Pgt$, $\Pe\Pgm$)             & 9:59                             & 760 \\
 3 ($\Pgm\Pgt$, $\Pe\Pgt$, $\Pe\Pgm$)             & 9:54                             & 767 \\
%4 ($\Pgm\Pgt$, $\Pe\Pgt$, $\Pe\Pgm$, $\Pgt\Pgt$) & 10:46                            & 712 \\
 4 ($\Pgm\Pgt$, $\Pe\Pgt$, $\Pe\Pgm$, $\Pgt\Pgt$) & 10:01                            & 758 \\
\bottomrule
\end{tabular}
\caption[Measurement of the runtime of the baseline $\PH\to\Pgt\Pgt$ analysis.]{Measurement of the runtime of the baseline $\PH\to\Pgt\Pgt$ analysis.
The analysis is performed on 10 files with a total size of 6.1~GB containing 452229 simulated $\PZ+\text{jet}$ events.
The measurement is performed with different numbers of pipelines, where each pipeline implements the analysis of one sub-channel of the analysis.
The values are compared with the execution without any pipelines and without global processors.
It is clearly visible, that the time needed for reading the inputs amounts to a large part of the overall runtime and that the simultaneous execution of multiple pipelines can reduce the runtime significantly compared to the case where different configuration are performed sequentially including separate IO operations.}
\label{table_artus_runtime_comparison}
\end{table}

>>>>

\usepackage{booktabs}
\usepackage{hepnames}
\usepackage{siunitx}
\usepackage{url}
\usepackage{palatino}
\usepackage[hidelinks,bookmarks=true,plainpages=false]{hyperref}
\usepackage[colorinlistoftodos,prependcaption]{todonotes}

@@

\begin{frontmatter}
\title{\artus{} -- Framework for Event-based Data Analysis in High Energy Physics}

\author{Joram Berger}
\author{Fabio Colombo}
\author{Raphael Friese}
\author{Dominik Haitz}
\author{Thomas Hauth}
\author{Thomas M\"uller}
\author{G\"unter Quast}
\author{Georg Sieber}
\address{Karlsruhe Institute of Technology, Wolfgang-Gaede-Stra\ss{}e 1, D-76131 Karlsruhe}

@@

\section{Typical Workflow of High Energy Physics Analyses}
\label{section_artus_analysis_workflow}
% This section should contain a short description of a HEP/CMS-typical workflow.
% Especially common problems and challenges should be highlighted.
% From these problems, the design requirements for Artus are derived.
The workflow of the experiments collecting data at the Large Hadron Collider (LHC) at CERN is taken as an example to illustrate the organization and needs of a modern High Energy Physics (HEP) analysis.
Nonetheless, equal or very similar requirements are also present for HEP experiments not located at the LHC.

@@

The reconstruction of the raw data coming from the experiments is performed on large computing centers organized in the Worldwide LHC Computing Grid (WLCG)~\cite{wlcg}.
The resulting datasets contain event information and high-level physics objects and are in the order of \SI{100}{TB} in size.

The data which are relevant for a specific analysis are selected in a step usually referred to as \textit{skimming}.
The output of the skimming process is an analysis-dependent file, usually in an n-tuple format, and small enough to be stored on local file servers, often owned and operated by a single institute.
The size of these files ranges from several \SI{10}{GB} to few~TB.

@@

\todo[inline]{I would not put the remaining part of this chapter here.
It should be either part of the requirements or part of the summary what Artus does achieve; but in this technical part in between it feels misplaced
-> Should we put this part into a dedicated section directly after this one?
I think the idea was to explain how certain design decisions, e.g.\ pipelines, achieve the \artus design goals.
Therefore it should not be too far from the technical descriptions.
}
This structure extends the established concepts of existing frameworks~\cite{cmssw} to the following four main advantages:
\begin{description}
\item[Modular structure] Analysis steps are implemented in processors, breaking down the complete analysis in well-defined and easily understandable and writable parts.
\item[Re-utilization of code] Processors can be shared among different analysis groups.
Every user contributes to the testing of existing code and improves the confidence in this code.
\item[Configurability] Each module uses settings which are read from the the configuration.
A separate sub-set of the configuration is created for each pipeline.
\item[High performance] The concept of pipelines allows to analyze the data with different configurations at the same time, while the data is read in only once and common processing steps are shared in the global pipeline.
Therefore, the cost-intensive input operations are reduced to a minimum.
\end{description}

@@

%time HiggsToTauTauAnalysis.py @HiggsAnalysis/KITHiggsToTauTau/data/ArtusWrapperConfigs/Sync13TeV.cfg -i HiggsAnalysis/KITHiggsToTauTau/data/Samples/NAF_sample_DYJetsToLLM50_RunIISpring15DR74_Asympt25ns_13TeV_MINIAOD_amcatnloFXFX-pythia8_recent.txt -f 10
\begin{table}[h!]
\centering
\begin{tabular}{lrr}
\toprule
\textbf{Pipelines}                                & \textbf{Runtime / \si{min}} & \textbf{Event Rate / \si{s^{-1}} } \\ \midrule
%0                                                & 5:53                             & 1296 \\
 0                                                & 5:12                             & 1470 \\ \midrule
%\em                                              & 6:27                             & 1181 \\
 1 ($\Pe\Pe$)                                     & 6:31                             & 1174 \\
%1 ($\Pe\Pgt$)                                    & 7:03                             & 1084 \\
 1 ($\Pe\Pgt$)                                    & 7:26                             & 1028 \\
%1 ($\Pgm\Pgt$)                                   & 7:14                             & 1054 \\
 1 ($\Pgm\Pgt$)                                   & 7:14                             & 1052 \\
%1 ($\Pgt\Pgt$)                                   & 6:14                             & 1224 \\
 1 ($\Pgt\Pgt$)                                   & 6:07                             & 1247 \\ \midrule
%2 ($\Pgm\Pgt$, $\Pe\Pgt$)                        & 8:33                             & 893 \\
 2 ($\Pgm\Pgt$, $\Pe\Pgt$)                        & 8:32                             & 896 \\
%3 ($\Pgm\Pgt$, $\Pe\Pgt$, $\Pe\Pgm$)             & 9:59                             & 760 \\
 3 ($\Pgm\Pgt$, $\Pe\Pgt$, $\Pe\Pgm$)             & 9:54                             & 767 \\
%4 ($\Pgm\Pgt$, $\Pe\Pgt$, $\Pe\Pgm$, $\Pgt\Pgt$) & 10:46                            & 712 \\
 4 ($\Pgm\Pgt$, $\Pe\Pgt$, $\Pe\Pgm$, $\Pgt\Pgt$) & 10:01                            & 758 \\
\bottomrule
\end{tabular}
\caption[Measurement of the runtime of the baseline $\PH\to\Pgt\Pgt$ analysis.]{Measurement of the runtime of the baseline $\PH\to\Pgt\Pgt$ analysis.
The analysis is performed on 10 files with a total size of 6.1~GB containing 452229 simulated $\PZ+\text{jet}$ events.
The measurement is performed with different numbers of pipelines, where each pipeline implements the analysis of one sub-channel of the analysis.
The values are compared with the execution without any pipelines and without global processors.
It is clearly visible, that the time needed for reading the inputs amounts to a large part of the overall runtime and that the simultaneous execution of multiple pipelines can reduce the runtime significantly compared to the case where different configuration are performed sequentially including separate IO operations.}
\label{table_artus_runtime_comparison}
\end{table}
